\section*{Projectile Simulation}

There is a projectile which is launched with
\[
\vec{r_0} = (x_0, y_0)
\]
\[
\vec{v_0} = (v_{0x}, v_{0y})
\]

The projectile is affected by gravity:
\[
\vec{F_g} = (0, -mg)
\]

The projectile is also affected by air resistance:
\[
\vec{F_r} = (-k v_x, -k v_y)
\]

There are some lines that the projectile can collide with. Each line is represented by two points:
\[
p_i = (\vec{p_{i1}}, \vec{p_{i2}})
\]

And there is constraint that
\[
\forall i, \vec{d_i} \parallel \hat{x} \lor \vec{d_i} \parallel \hat{y}
\]
\[
\forall i, \vec{p}_{i2} = \vec{p}_{(i+1)1}
\]

where $\vec{d_i} = \vec{p_{i2}} - \vec{p_{i1}}$.


\vspace{3em}
\hrule
\vspace{2em}

\subsection*{Proof that the projectile cannot collide with a vertical edge}

We consider a projectile with initial position
\[
\vec r_0 = (x_0, y_0)
\]
and initial velocity
\[
\vec v_0 = (v_{0x}, v_{0y})
\]

The forces acting on the projectile are gravity
\[
\vec F_g = (0, -mg)
\]
and linear air resistance
\[
\vec F_r = (-k v_x, -k v_y)
\]

We prove that the projectile cannot collide with any edge parallel to the $y$ axis.

\subsection*{Step 1: Equation of motion in the $x$ direction}

By Newton's second law,

\[
m \frac{dv_x}{dt} = -k v_x
\]

This is a separable differential equation:

\[
\frac{dv_x}{v_x} = -\frac{k}{m} dt
\]

Integrating:

\[
\ln |v_x| = -\frac{k}{m} t + C
\]

Thus,

\[
v_x(t) = v_{0x} e^{-\frac{k}{m}t}
\]

\subsection*{Step 2: Position function $x(t)$}

Integrating velocity:

\[
x(t) = x_0 + \int_0^t v_x(s) ds
\]

\[
x(t) =
x_0 + \frac{m}{k} v_{0x}
\left(1 - e^{-\frac{k}{m}t}\right)
\]

\subsection*{Step 3: Monotonicity of $x(t)$}

Since

\[
e^{-\frac{k}{m}t} > 0 \quad \forall t
\]

we have

\[
v_x(t) = v_{0x} e^{-\frac{k}{m}t}
\]

Therefore the sign of $v_x(t)$ is constant:

\[
\text{sign}(v_x(t)) = \text{sign}(v_{0x})
\]

Thus,

\[
\frac{dx}{dt} = v_x(t)
\]

never changes sign.

Therefore,

\[
x(t)
\]

is a strictly monotonic function.

\subsection*{Step 4: Vertical edge definition}

A vertical edge is defined by

\[
x = c
\]

for some constant $c$.

A collision would require two distinct times $t_1 \neq t_2$ such that

\[
x(t_1) = x(t_2) = c
\]

\subsection*{Step 5: Application of the Intermediate Value Theorem}

The function $x(t)$ is continuous.

Since it is strictly monotonic, it is injective.

Therefore,

\[
x(t_1) = x(t_2)
\implies
t_1 = t_2
\]

Thus, there cannot exist two distinct times where the projectile has the same $x$ coordinate.

\subsection*{Conclusion}

Since a collision with a vertical edge would require the projectile to reach the same $x$ coordinate more than once, and since $x(t)$ is strictly monotonic, this is impossible.

Therefore,

\[
\boxed{
\text{The projectile cannot collide with any edge parallel to the $y$ axis.}
}
\]

\vspace{3em}
\hrule
\vspace{2em}

\section*{Numerical Integration Convergence}

In physics simulations, we often use simple update rules to approximate continuous motion. This section explains that as the time step $dt \to 0$, these numerical updates converge to the true analytical physical solutions.

\subsection*{Case 1: Drag Force (Velocity Convergence)}

\textbf{Mathematical Model:}
Consider a particle subject to a drag force proportional to its velocity:
\[
F = -k v \implies m \frac{dv}{dt} = -k v \implies \frac{dv}{dt} = -\frac{k}{m} v
\]
The analytical solution is $v(t) = v_0 e^{-(k/m)t}$.

\textbf{Discretized Update Rule:}
In a simulation, we update the velocity using the rule:
\[
v += -(k/m) v \cdot dt \implies v_{n+1} = v_n \left(1 - \frac{k}{m} dt\right)
\]

\textbf{Convergence Explanation:}
After $n$ steps ($t = n \cdot dt$), the velocity is:
\[
v_n = v_0 \left(1 - \frac{k}{m} dt\right)^n = v_0 \left(1 - \frac{k}{m} \frac{t}{n}\right)^n
\]
Using the definition of the exponential function $\lim_{n \to \infty} (1 - \frac{a}{n})^n = e^{-a}$:
\[
\lim_{dt \to 0} v_n = v_0 e^{-(k/m)t}
\]
Thus, the simulation result converges to the actual physical solution as $dt \to 0$.

\subsection*{Case 2: Integrating Velocity to Position}

\textbf{Mathematical Model:}
Given a velocity function $v(t)$, the continuous equation is $dx/dt = v(t)$, which yields:
\[
x(t) = x_0 + \int_0^t v(s) ds
\]

\textbf{Discretized Update Rule:}
The simulation repeatedly applies:
\[
x += v \cdot dt \implies x_{n+1} = x_n + v(t_n) dt
\]

\textbf{Convergence Explanation:}
Summing the updates yields:
\[
x_n = x_0 + \sum_{k=0}^{n-1} v(t_k) dt
\]
This is exactly the \textbf{Riemann Sum} of the velocity function. As $dt \to 0$, the Riemann sum converges to the definite integral:
\[
\lim_{dt \to 0} x_n = x_0 + \int_0^t v(s) ds = x(t)
\]

\subsection*{Case 3: Full Trajectory (Acceleration to Position)}

\textbf{Mathematical Model:}
Given an acceleration $a(t)$, the system is:
\[
\frac{dv}{dt} = a(t), \quad \frac{dx}{dt} = v(t)
\]

\textbf{Discretized Update Rule:}
The simulation updates both velocity and position:
\[
v += a \cdot dt, \quad x += v \cdot dt
\]

\textbf{Convergence Explanation:}
As shown in Case 2, the velocity update converges to $v(t) = v_0 + \int a(t) dt$. Then, the position update, which uses these converged velocity values, also converges to the integral of $v(t)$. Therefore, the entire chain $a \to v \to x$ converges to the physical trajectory as $dt \to 0$.
